The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley \revision{Radio Observatory} Long Wavelength Array

### 2018-04-26 paper.tex
update the ionospheric text and other minor edits

Current 21-cm cosmology experiments can be broadly separated into two classes: global signal
experiments that aim to detect the spectral signature of the cosmologically redshifted 21-cm
transition after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21-cm perturbations.  Ongoing global signal experiments include EDGES
\citep{2010Natur.468..796B, 2017ApJ...847...64M}, LEDA \citep{2017arXiv170909313P}, BIGHORNS
\citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2
\citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \revision{\citet{2017MNRAS.464.4995M}} as
\revision{\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.6}\,.
\end{equation}}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally, a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.

@@

In the context of $m$-mode analysis a transit telescope is any interferometer for which the response
pattern of the individual elements does not change with respect to time. This may be an
interferometer like the OVRO-LWA where the correlation elements are fixed dipoles, but it may also
be an interferometer like LOFAR or the MWA if the steerable beams are held in a fixed position (not
necessarily at zenith). The interferometer also does not necessarily have to be homogeneous.
Heterogeneous arrays composed of several different types of antennas are allowed as long as care is
taken to generalize Equation~\ref{eq:basic-imaging} for a heterogeneous array.

@@

If the baselines have a noise covariance matrix $\b N \neq n\b I$ for some scalar $n$ (eg. the
interferometer is inhomogeneous), then the observer should minimize $(\b v-\b B\b a)^*\b N^{-1}(\b
v-\b B\b a) + \varepsilon\|\b a\|^2$. The matrix $\b N$ is used to weight the measurements such that
\begin{equation}
    \b{\hat a}_\text{min variance} = (\b B^*\b N^{-1}\b B + \varepsilon\b I)^{-1}
        \b B\b N^{-1}\b v\,.
\end{equation}

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/scintillation-refraction/scintillation-refraction}
    \caption{
        Panels (a) and (b) show the measured apparent flux of Cyg~A and Cas~A at 36.528 MHz (red
        points) and 73.152 MHz (blue points) as a function of time over the observing period. The
        solid black curves show the expected flux computed using the empirical beam model fits. The
        thermal noise contribution to each point is about 50 Jy.  Cyg~A is occulted by the White
        Mountains when it is low on the horizon to the east.
        Panels (c) and (d) These panels show the measured position offset of Cyg~A and Cas~A
        relative to their true astronomical positions at 36.528 MHz (red line) and 73.152 MHz (blue
        line).
    }
    \label{fig:scintillation}
\end{figure*}

@@

\revision{To achieve this dynamic range at low frequencies, it is important to account for
propagation effects through the ionosphere. In the weak scattering regime ($b_\text{max} \ll 2\pi
r_\text{diff}$, where $b_\text{max}$ is the maximum baseline length, and $r_\text{diff}$ is the
diffractive scale of the ionosphere), fluctuations within the ionosphere contribute amplitude and
phase scintillations that can be described by a direction-dependent complex-gain calibration. This
justifies the use of ``peeling'', which incorporates a direction-dependent calibration to subtract
sources in the presence of ionospheric scintillation.  \citep[eg.][]{2008ISTSP...2..707M,
2015MNRAS.449.2668S}.}

\revision{In the strong scattering regime ($b_\text{max} \gtrsim 2\pi r_\text{diff}$), the image of
a point source can ``break apart'' into multiple images or speckles \citep{2015MNRAS.453..925V}.
Attempting to peel a source in the strong scattering regime will lead to source subtraction
artifacts in the final sky map.  \citet{2016RaSc...51..927M} measured that from the location of
LOFAR at 150 MHz, the diffractive scale of the ionosphere is $>5$ km 90\% of the time. This implies
that at 73 MHz, the diffractive scale is typically $>2$ km, and at 36 MHz, the diffractive scale is
typically $>1$ km. The maximum baseline length of the OVRO-LWA is $\sim 1.5$ km, and therefore we do
not generally expect to see strong scattering from the ionosphere.  Furthermore, ionospheric
conditions during the observing period were mild (see \S\ref{sec:ionosphere-conditions}), but we
still observe scintillation and refractive-offset events on the timescale of a single integration
(13 seconds, see Figure~\ref{fig:scintillation}). Consequently we peeled Cyg~A and Cas~A from the
dataset using a new solution for each integration.}

\revision{In addition, the largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular
scale of Cyg~A is $\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines at
73~MHz, the OVRO-LWA marginally resolves both sources. A resolved source model is needed for both
sources. We fit a self-consistent resolved source model to each source. This is performed by
minimizing the variance within an aperture located on each source after peeling. By phasing up a
large number of integrations before imaging (at least 1 hour), it is possible to smear out the
contribution of the rest of the sky.  We then use a non-linear optimization routine (NLopt Sbplx)
\citep{nlopt, sbplx} to vary the parameters in a source model until the variance within the aperture
is minimized. Cyg~A is modeled with 2 Gaussian components, while Cas~A is modeled with 3~Gaussian
components.  Ultimately, these multi-component models are used to peel Cyg~A and Cas~A, but residual
errors from this model and from the ionosphere (particularly while these sources are at low
elevations) contribute residual artifacts that are largely localized to within 1$^\circ$ of each
source.}

@@

In order to understand the effect of polarization leakage, we simulated a point source with 10\%
polarization in Stokes-$Q$ at the location of Cas A.  The simulated visibilities were computed using
the measured beams for $x$- and $y$-dipoles. Because the amplitude of the two beams are not equal in
every direction on the sky, this introduces a direction-dependent leakage of Stokes-$Q$ into
Stokes-$I$. At 73.152~MHz, this leakage is $\lesssim5\%$ above $15^\circ$ elevation, but rapidly
rises to $\gtrsim50\%$ at lower elevations. \citet{2015JAI.....450004O} report similar polarization
leakage measurements with the LWA1.  Cas A is a circumpolar source and spends about seven hours
every day skirting the horizon where the polarization leakage is large, so by placing the simulated
source at the location of Cas A, we are engineering a situation where the polarization leakage from
Stokes-$Q$ into Stokes-$I$ will be large. However, the impact on the unpolarized $m$-mode analysis
maps is mild, amounting to a 0.5\% error in the flux of the source with no measurable effect on the
PSF.

>>>>

Current 21-cm cosmology experiments can be broadly separated into two classes: global signal
experiments that aim to detect the spectral signature of the cosmologically redshifted 21-cm
transition after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21-cm perturbations.  Ongoing global signal experiments include EDGES
\citep{2017ApJ...847...64M}, LEDA \citep{2017arXiv170909313P}, BIGHORNS \citep{2015PASA...32....4S},
SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2 \citep{2017arXiv170306647S}.  Ongoing power spectrum
experiments include PAPER/HERA \citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR
\citep{2017ApJ...838...65P}, and the MWA \citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}. Recently,
EDGES reported the first detection of 21-cm absorption in the globally averaged sky-signal
\citep{2018Natur.555...67B}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \revision{\citet{2017MNRAS.464.4995M}} as
\revision{\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.6}\,.
\end{equation}}
Therefore, experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P} and \citealt{2013MNRAS.429..165C}) rely on the
assumption that the foreground emission is spectrally smooth. The low-frequency radio sky is
composed of several components: galactic synchrotron emission, supernova remnants, radio galaxies,
free-free emission and absorption from \ion{H}{2} regions, and a confusing background of radio
sources.  Ideally, a foreground removal strategy should be informed by the measured spatial
structure and frequency spectrum of all foreground components. For instance, CMB experiments
typically construct several maps at several frequencies to enable component separation.  At low
frequencies, this possibility is limited by the availability of suitable high-fidelity sky maps on
angular scales ranging from tens of degrees to arcminutes.

@@

In the context of $m$-mode analysis, a transit telescope is any interferometer for which the
response pattern of the individual elements does not change with respect to time. This may be an
interferometer like the OVRO-LWA where the correlation elements are fixed dipoles, but it may also
be an interferometer like LOFAR or the MWA if the steerable beams are held in a fixed position (not
necessarily at zenith). The interferometer also does not necessarily have to be homogeneous.
Heterogeneous arrays composed of several different types of antennas are allowed as long as care is
taken to generalize Equation~\ref{eq:basic-imaging} for a heterogeneous array.

@@

If the measured $m$-modes have a noise covariance matrix $\b N \neq n\b I$ for some scalar $n$ (eg.
the interferometer is inhomogeneous), then the observer should minimize $(\b v-\b B\b a)^*\b
N^{-1}(\b v-\b B\b a) + \varepsilon\|\b a\|^2$. The noise covariance matrix $\b N$ is used to weight
the measurements such that
\begin{equation}
    \b{\hat a}_\text{min variance} = (\b B^*\b N^{-1}\b B + \varepsilon\b I)^{-1}
        \b B\b N^{-1}\b v\,.
\end{equation}

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/scintillation-refraction/scintillation-refraction}
    \caption{
        Panels (a) and (b) show the measured apparent flux of Cyg~A and Cas~A at 36.528 MHz (red
        points) and 73.152 MHz (blue points) as a function of time over the observing period. The
        solid black curves show the expected flux computed using the empirical beam model fits. The
        thermal noise contribution to each point is about 50 Jy.  Cyg~A is occulted by the White
        Mountains when it is low on the horizon to the east.
        Panels (c) and (d) show the measured position offset of Cyg~A and Cas~A relative to their
        true astronomical positions at 36.528 MHz (red line) and 73.152 MHz (blue line).
    }
    \label{fig:scintillation}
\end{figure*}

@@

\revision{
    To achieve this dynamic range at low frequencies, it is important to account for propagation
    effects through the ionosphere. In the weak scattering regime ($r_\text{diff} \gg r_\text{f}$,
    where $r_\text{diff}$ is the diffractive scale of the ionosphere, $r_\text{f} =
    \sqrt{\lambda D / 2\pi}$ is the Fresnel scale, $\lambda$ is the wavelength, and $D$ is the
    distance to the ionosphere), fluctuations within the ionosphere contribute
    amplitude and phase scintillations that can be described by a direction-dependent complex-gain
    calibration. This justifies the use of ``peeling'', which incorporates a direction-dependent
    calibration to subtract sources in the presence of ionospheric scintillation.
    \citep[eg.][]{2008ISTSP...2..707M, 2015MNRAS.449.2668S}.
}

\revision{
    In the strong scattering regime ($r_\text{diff} \lesssim r_\text{f}$), the image of a point
    source can ``break apart'' into multiple images or speckles \citep{2015MNRAS.453..925V}.
    Attempting to peel a source in the strong scattering regime will lead to source subtraction
    artifacts in the final sky map.  \citet{2016RaSc...51..927M} measured that from the location of
    LOFAR at 150 MHz, the diffractive scale of the ionosphere is $>5$ km 90\% of the time. This
    implies that at 73 MHz, the diffractive scale is typically $>2$ km, and at 36 MHz, the
    diffractive scale is typically $>1$ km. These limits are comparable to the Fresnel scale for the
    OVRO-LWA (ie. $r_\text{diff} > r_\text{f}$), and therefore we do not generally expect to see
    strong scattering from the ionosphere.  Ionospheric conditions during the observing period were
    mild (see \S\ref{sec:ionosphere-conditions}). However, we do observe scintillation and
    refractive-offset events on the timescale of a single integration (13 seconds, see
    Figure~\ref{fig:scintillation}). Consequently we peeled Cyg~A and Cas~A from the dataset using a
    new solution for each integration.
}

\revision{
    In addition, the largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular scale
    of Cyg~A is $\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines at
    73~MHz, the OVRO-LWA marginally resolves both sources. A resolved source model is needed for
    both sources. We fit a self-consistent resolved source model to each source. This is performed
    by minimizing the variance within an aperture located on each source after peeling. By phasing
    up a large number of integrations before imaging (at least 1 hour), it is possible to smear out
    the contribution of the rest of the sky.  We then use a non-linear optimization routine (NLopt
    Sbplx) \citep{nlopt, sbplx} to vary the parameters in a source model until the variance within
    the aperture is minimized. Cyg~A is modeled with 2 Gaussian components, while Cas~A is modeled
    with 3~Gaussian components.  Ultimately, these multi-component models are used to peel Cyg~A and
    Cas~A, but residual errors from this model and from the ionosphere (particularly while these
    sources are at low elevations) contribute residual artifacts that are largely localized to
    within 1$^\circ$ of each source.
}

@@

In order to understand the effect of polarization leakage, we simulated a point source with 10\%
polarization in Stokes-$Q$ at the location of Cas A.  The simulated visibilities were computed using
the measured beams for $x$- and $y$-dipoles. Because the amplitude of the two beams are not equal in
every direction on the sky, this introduces a direction-dependent leakage of Stokes-$Q$ into
Stokes-$I$. At 73.152~MHz, this leakage is $\lesssim5\%$ above $15^\circ$ elevation, but rapidly
rises to $\gtrsim50\%$ at lower elevations. \citet{2015JAI.....450004O} report similar polarization
leakage measurements with the LWA1.  Cas A is a circumpolar source and spends about seven hours
every day skirting the horizon where the polarization leakage is large, so by placing the simulated
source at the location of Cas A, we are engineering a situation where the polarization leakage from
Stokes-$Q$ into Stokes-$I$ will be large. However, the impact on the unpolarized $m$-mode analysis
maps is mild, amounting to a 0.5\% error in the flux of the source.
